---
title: "Analyzing Yelp Data"
output: pdf_document
---

```{r}
clean_data = read.csv("/Users/jeekeatwah/Downloads/Stats 101C Files/Data_Clean")
```

```{r}
# analyzing positive to negative words ratio
postoneg_1 = clean_data[clean_data$Ratio.of.Positive.to.Negative.Words == 1, ]
nrow(postoneg_1)
table(postoneg_1$Star)
```

```{r}
# positive to negative words ratio > 0.5
postoneg_0.5 = clean_data[clean_data$Ratio.of.Positive.to.Negative.Words > 0.5, ]
nrow(postoneg_0.5)
table(postoneg_0.5$Star)
```
Ratio > 0.5, higher likelihood of a higher star from a user

```{r}
table(postoneg_0.5$Bus_Ave_Star)
```
Ratio > 0.5, higher likelihood of having a higher business average rating

```{r}
# stars >= 4
starg4 = clean_data[clean_data$Star >= 4, ]
hist(starg4$Users_Ave_Star)
```
For individual reviews greater than or equal to 4 stars, the user's average star rating is not a factor


Since the dataset only contains users who have more than 150 total reviews, their ratings are, on average, pretty close to the actual average business rating
```{r}
stardiff_mean = mean(clean_data$Star - clean_data$Bus_Ave_Star)
stardiff_mean
```

The higher the popularity of a user's reviews (useful, cool, funny), the closer the user's rating to the actual business rating
```{r}
useful_median = median(clean_data$User_Useful_count)
funny_upmedian = median(clean_data$User_Funny_count)
cool_upmedian = median(clean_data$User_Cool_count)
```

```{r}
user_useful = clean_data[clean_data$User_Useful_count > useful_median, ]
(stardiff_mean - mean(user_useful$Star - user_useful$Bus_Ave_Star)) / stardiff_mean
```
Users that are more "useful" are 13.4% closer to the actual business rating than the overall mean

```{r}
user_funny = clean_data[clean_data$User_Useful_count > funny_upmedian, ]
(stardiff_mean - mean(user_funny$Star - user_funny$Bus_Ave_Star)) / stardiff_mean
```
Users that are more "funny" are 15.7% closer to the actual business rating than the overall mean

```{r}
user_cool = clean_data[clean_data$User_Useful_count > cool_upmedian, ]
(stardiff_mean - mean(user_cool$Star - user_cool$Bus_Ave_Star)) / stardiff_mean
```
Users that are more "cool" are 20.1% closer to the actual business rating than the overall mean